### 2020-07-07 paper.Rmd
misc minor

@@

- It is often important to divide a set of elements into similar 
subgroups
- This task can be automized and open source solutions exist 
- Anticlustering = opposite of clustering, does that
- Objectives based on pairwise distances exist, but k-means also exists 
- k-means = optimizes similarity wrt to mean, but not other distribution
parameters, in particular not wrt to the variance
- here, I present an improved objective function for k-means 
anticlustering, which optimizes similarity wrt to means and variance / 
standard deviation at the same time

@@

plot(df, pch = 4, col = "darkgrey", las = 1)
# Illustrate best partition wrt simililarity in means
points(
  all_objectives_mean[which.min(all_objectives_mean)],
  all_objectives_var[which.min(all_objectives_mean)],
  cex = 2, col = "#DF536B", pch = 16
)

@@

>>>>

bibliography      : ["lit.bib"]

@@

Oftentimes researchers have to divide a set of elements different, but 
equivalent groups. For example, this problem arises when 
splitting a test into parts of equal difficulty [@gierl2017], when 
assigning students to work groups [@baker2002], or when assigning 
stimuli to experimental conditions [@lahl2006]. Solving such problems 
manually is a tedious and time-consuming task and the similarity of 
manually established groups is usually improvable. Fortunately, these 
problems can be formalized mathematically [e.g., @brusco2019; 
@spath1986] and by now, accessible open source solutions for tackling 
this automatically exist [@papenberg2020].

Forming similar groups is also known as the *anticlustering* problem. 
Anticlustering reverses the logic of its better known twin, cluster 
analysis [@rokach2005]. In cluster analysis, groups are sought that are 
internally homogenous, whereas there is heterogeneity between groups 
[@steinley2006]. Anticlustering instead strives for high between-group 
similarity high within-group heterogeneity. Usually, cluster analysis 
aims to uncover "true" latent structure contained in a data set, whereas 
anticlustering imposes an "unnatural" division of the data.

One well-known criterion for anticlustering reverses the k-means 
clustering method by maximizing the within-group variation, or 
*variance* [@jain2010; @steinley2006; @spath1986; @valev1998; 
@valev1983]. Späth and Valev independently coined the term 
anticlustering specifically to denote the reversal of classical 
k-means.

--- k-means

--- dissimilarity data vs. rectangular data

Although forming groups with high between-group similarity is often 
equivalent to establishing high within-group diversity, researchers 
usually only wish to obtain one of these goals. In experimental 
psychology, sets of stimuli should be as similar as possible; within-set 
heterogeneity is only an unimportant side-product. In this paper, I 
address such problems that require high between-group similarity. I 
present a novel extension to the k-means objective function, 
within-group sum-of-squares that improves its capacity to form similar 
groups [@steinley2006], but is no longer aims to maximize heterogeneity.

Ironically, even though maximizing the k-means objective maximizes 
similarity with regard to the means of the

- here, I present an improved objective function for k-means 
anticlustering, which optimizes similarity wrt to means and variance / 
standard deviation at the same time

@@

plot(df, pch = 4, col = rainbow(length(partitions)), las = 1)
# Illustrate best partition wrt simililarity in means
points(
  all_objectives_mean[which.min(all_objectives_mean)],
  all_objectives_var[which.min(all_objectives_mean)],
  cex = 2, col = "#DF536B", pch = 16
)

@@

# Combined approach better than k-variance alone

- Often, maximizing k-means-variance yields more similar variances 
than k-variance alone. Why is that? I guess: structure in the solution 
space, local maxima are real. When also considering means, you more 
often escape bad local maxima?
